On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-01-13 MS.tex
Substantial edits to text
- still work to do merging theoretical section with hypotheses

\title{On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted}
\author{Anais Gibert\textasteriskcentered, Emma F. Gray, Mark Westoby, Ian J. Wright, Daniel S. Falster}
\affiliation{Biological Sciences, Macquarie University NSW 2109, Australia\\
\textasteriskcentered Email for correspondence: \texttt{anais.gibert@gmail.com}\\
Word count: ~5538 words (without supplementary material, Fig and ref)}
\date{}

@@

\mstype{Research Article}
\runninghead{Influence of traits on growth changes with size}
\keywords{absolute growth rate, maximum height, photosynthetic rate, ontogenetic stage, functional trait, growth strategy, relative growth rate, seed mass, specific leaf area, wood density}

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary 268/350}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate has often been seen as a central element of its ecological strategy, and as ultimately linked to its traits. Yet the literature is inconsistent about the empirical correlation between functional traits and growth, casting doubt on the capacity of traits to predict growth.

\item We propose that traits should influence growth in a way that depends on the size of individual plants. We outline mechanisms and hypotheses based on new theory, and test these predictions in tree species for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate) using a meta-analysis of 112 studies ($>$ 500 correlations). To capture the effects of size, we tested for a shift in the direction of correlation between growth rates and traits across three plant stages: seedling, sapling and adult.

\item Results were consistent with predictions. Specific leaf area was correlated with relative growth rate (RGR) in seedling but not in adult plants. Correlations of RGR with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with RGR only in one plant stage category: seedlings and adults, respectively. Despite the vast litterature on trait and growth, coverage was limited for wood density, seed mass, maximum height and assimilation rate, and additional studies would be helpful for some trait-growth-stage relationships.

@@

The phrase ``plant ecological strategy'' refers to how species face common challenges of acquiring sufficient water, nutrients and light for growth, and then ultimately replacing themselves with offspring. An idea that species strategy differences might involve fast versus slow growth goes back at least to \citet{Grime:1975gr}, who measured a wide spread of relative growth rates of seedlings during early exponential growth under favourable conditions. Seedling potential relative growth rate (RGR) has continued to be used as a major strategy indicator within the CSR (Competitor - Stress tolerator - Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling RGR into different components showed that specific leaf area (SLA = 1/leaf mass per area LMA) was typically the strongest source of variation between species \citep{Poorter:1989tx, Rees:2010gk}. From the 1990s a new style of ``trait ecology'' arose whereby measurable traits including SLA were used directly as strategy axes \citep{Westoby:2002ft}. An approach via measurable traits made broad comparisons possible across continents and latitudes and thousands of species. Since then, traits have been seen as constructive approaches to understanding plant growth strategies \citep{Grime:1977kc,Chapin:1980gz}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

Currently, two main spectra of variation are widely cited as underpinning differences between species in vegetative growth rates. The ``leaf economic spectrum'' \citep{Wright:2004jb} and the ``wood economic spectrum'' \citep{Chave:2009iy} reflect tissue construction costs (for leaf area and for wood volume respectively), trading off with rates of tissue turnover or mortality. It has been widely expected that species with low tissue construction cost will typically have fast growth rates, at least in favourable physical environments \citep[e.g.][]{MullerLandau:2004dc,Poorter:2008iu,Chave:2009iy,Larjavaara:2010bn,Iida:2012jb,Paine:2015df}. Yet, recently a number of empirical studies indicate the correlation between traits and growth is not as consistent as that. Consider SLA as an illustration. Species with high SLA can deploy more canopy area for a given dry mass investment in leaves \citep{Poorter:1999wd, Reich:1992wm}. In addition, SLA is correlated with specific photosynthetic capacity per unit of leaf mass and high nitrogen (N) or phosphorus concentrations in the leaf \citep{Wright:2004jb}. For seedling relative growth rates under favourable conditions, high SLA has repeatedly been found to be a strong predictor regardless of vegetation type or growth form \citep{Lambers:1992bj,Reich:1992wm,Grime:1997wm,Poorter:1999wd,Wright:1999ds}. Yet for adult plants, the correlation has mostly been much weaker or absent \citep{coomes_comparison_1998,Poorter:2008iu,Aiba:2009ft,Easdale:2009gv,Wright:2010tp}. The discrepancy between theoretical expectation and empirical results has led some to question whether traits are even useful for understanding growth \citep{Wright:2010tp, Paine:2015df}.

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Size has long been known to play a central role in organismal design and function \citep{Vogel:1988ux, Vogel:2003wb}, including that of trees \citep{Farnsworth:1995im, Givnish:1995ta}. Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii}. At larger sizes the low-cost deployment benefits of high SLA may be offset or even outweighed by the high costs of rapid leaf turnover (see section ``Theoretical expectations''). While recent empirical studies supported the idea of a variation in the relationships between traits and growth rates across plant sizes (\citealt{Iida:2014ep, Iida:2014hq}), the trend and mechanisms behind this variation remain unclear.

In this paper, we used a meta-analytic approach on five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass) to investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item  Do the observed relationships conform to predictions (outlined in section ``Theoretical expectations'') derived from a recently developed mechanistic model \citep{Falster:2011ii}?
\end{enumerate}
We gathered correlation coefficients between traits and growth from studies that gave explicit information about the size or stage of the plants. While most studies on traits have not been concerned with effects of plant size, they have generally reported trait-growth correlations for identifiable stage classes. We used plant stage (i.e. seedling, sapling and adult) as a surrogate for the effects of plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.

Meta-analytic approach allows generalization by summarizing and integrating research results across studies, just like large data compilation commonly used in trait ecology. Yet, a specific interest of meta-analytic approach is also to identify key knowledge gaps. Here, we expected a vast literature on trait and growth to be available, but we were uncertain about the representation of all trait-stages combinations. Thus, a goal of our study was also to identify areas of research needed to improve our understanding of the link between trait and growth.

@@

Recent theoretical work \citep{Falster:2011ii} allows us to formulate hypotheses about relationships between traits and growth, based on a mathematical decomposition of plant growth rate. Specific predictions arise from considering the influence of traits on elements of these equations and how these effects may vary with size (see Appendix \ref{sec:growth} for mathematical explanations). This model extends earlier efforts to model growth as a function of traits \citep{Lambers:1992bj,cornelissen_seedling_1996,Wright:2000kw,Enquist:2007ek}. However, whereas earlier models focused solely on mass-based relative growth rate and did not predict a change in effect of traits with size, the revised equations do allow for these effects, while also making predictions for both absolute and relative growth rates in any of height, diameter or mass.

@@

Plants with high SLA can deploy more leaf area per unit of mass invested. However, high-SLA species also have shorter leaf lifespan and faster leaf turnover rate \citep{Wright:2004jb}. Whether plant growth increases with SLA depends on the relative magnitude of the leaf construction and leaf turnover effects of SLA (Table \ref{tab:trade-offs}). During early life, when leaf comprises a large part of the plant, decreasing leaf construction cost has a positive effect on growth rate. As plants increase in size and leaf mass fraction decreases, the benefit of cheaper leaf construction diminishes; thus the cost of replacing leaf turnover becomes more important.

@@

High WD increases the cost of building stem tissue and thus the cost of deploying additional leaf area (Table \ref{tab:trade-offs}), which should result in slower growth. Although expensive wood decreases the risk of mortality, this benefit is not reflected in growth outcomes. Thus increasing WD is predicted to decrease growth across the entire size range.

\textbf{H3: The relationship between Hmax and growth rate is expected to be absent in seedlings but to strengthen with increasing plant size.}

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do,Thomas-2011, Wenk:2014jz}). For seedlings, where all individuals are investing solely in growth, we expect no influence of Hmax on growth. As plants become larger, species with smaller Hmax are expected to be allocating more to reproduction and less to growth, because at any given size they are closer to their Hmax (Table \ref{tab:trade-offs}).

\textbf{H4: The relationship between seed mass and relative growth rate is negative in seedlings, but weakens with increasing size.}

The main effect of increased seed size is to produce a larger overall seedling. Under similar light conditions, relative growth rate decreases with size because the ratio of leaf area to total biomass decreases with increasing plant size. Thus, larger seedlings are predicted to have a lower relative growth rate during the earliest phases of growth \citep{Turnbull:2012ew}. However, among larger plants we expect the effect of this initial size to be minimal, and thus the correlation between seed size and growth rate should disappear.

\textbf{H5: The relationship between Aarea and growth rate is expected to be positive across all plant sizes, at least under high-light conditions}

@@

\textbf{Effect on absolute and relative growth rates}
Tree growth in diameter, height or mass can be expressed in either absolute (AGR) or relative terms (RGR). The choice of the growth measurement is generally confunded with the size of the plant; RGR in mass is mainly used for seedling, and AGR in diameter or height are mainly used for adult. According to the growth model, the predicted relationship between growth and seed mass in seedlings reverses if absolute AGR$_\textrm{mass}$ is used instead of RGR$_\textrm{mass}$ (see Appendix \ref{sec:growth}). The correlation with RGR$_\textrm{mass}$ is negative but becomes positive with AGR$_\textrm{mass}$. For SLA, WD, Hmax and Aarea the predicted relationships are not affected by the growth measurement used. However RGR and AGR are fundamentaly different for a lot of ecologists. So we adopted a two steps approach, first we were conservative by focusing exclusively on RGR, then we compared results with those of a more inclusive dataset with both RGR and AGR.

@@

We aimed to identify studies that measured the correlation between five prominent functional traits (SLA, WD, Aarea, Hmax and seed mass) and a growth measurement across species. Our literature search was conducted in two phases.

\subsubsection*{Phase 1 - Literature snapshot}\label{literature-snapshot}

\subsubsection*{Phase 2 - Systematic search}\label{systematic-search}

While Web of Science guarantees a relatively stable and reproducible search environment, with clearly defined lists of indexed journals, its searches are limited to meta-data (i.e. title, keywords and abstracts, \citealt{Beckmann:2012hn}). This limitation does not apply to Google Scholar, where all text that is electronically available on the Internet will be indexed. Thus, we used Google Scholar to perform a more systematic search of the literature for SLA and WD.

The Google Scholar database was searched using a range of Boolean search terms for SLA and WD: ((``SLA'' OR ``LMA'' OR ``specific leaf area'' OR ``leaf mass area'' OR ``wood density'' OR ``WD'' OR ``wood specific gravity'' OR ``stem density'' OR ``stem specific density'') AND (``growth rate'' OR ``relative growth rate'')) AND (``tree*'' OR ``woody*''). This keyword search in Google Scholar returned approx. 18900 papers, from which we took the first 750 items when ranked by ``relevance''.
These papers were examined in addition to those discovered in Phase 1. In addition, we gleaned data from other sources, using our own knowledge as researchers in the field (this included adding data from an in-review study by \citeauthor{Gleason:2016}). The complete list of studies where we inspected the full text is given in Appendix \ref{app:supp_info_files} in Supplementary material.

Each study had to meet some basic criteria to be included in our analysis. We restricted our meta-analysis to documents reporting: i) the relationship between growth and trait as a correlation coefficient, ii) the sample size and the number of species used to perform the correlation, and iii) the size/stage of plants. Ultimately data was included from 112 different articles (starred references in bibliography). Where available, we also recorded the geographic coordinates, number of sites, experiment type (greenhouse, field experiment, forest, shade-house etc.), growth measure (relative or absolute growth rate measured using plant diameter, height or mass), vegetation type, life form (tree, woody or mixed with herbaceous species etc.), biological scale (interspecific, intraspecific or intragenus), and experimental treatments (e.g. fertilization, shade, salt stress).  We asked the authors reporting a non-significant correlation for additional information about the specific correlation coefficient value when this was not provide. This allowed us to minimise publication bias via selective reporting of significant results within the published articles (e.g. a particular case of a ``failure to publish bias'' see \citealt{Jennions:2013ta}).

Overall, we recorded coefficient correlations for a list of 19 traits (Fig. \ref{fig:fig1}): the five for which we had prior expectations, plus 14 others commonly used in the trait literature. SLA, leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and Aarea are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. Seed mass and Hmax capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. WD and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here WD and SA/LA are also considered as indicators of stem construction cost \citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource distribution to leaf \citep{Lambers:1992bj}. According to the snapshot of the literature on Web of Science, 62\% of the literature for these 19 traits was captured by a search on SLA and WD.

>>>>

\title{On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted}
\author{Ana{\"i}s Gibert\textasteriskcentered, Emma F. Gray, Mark Westoby, Ian J. Wright, Daniel S. Falster}
\affiliation{Biological Sciences, Macquarie University NSW 2109, Australia\\
\textasteriskcentered Email for correspondence: \texttt{anais.gibert@gmail.com}\\
Word count: ~5538 words (without supplementary material, Fig and ref)}
\date{}

% Todo:
% Ditch abbrvos for SLA, wood density, Hmax?
% Fix plots
%   -- labels causing plot size to change, e.g. Fig 3d
%   -- only one y label for multi-panel plots
%   -- sample sizes skewiff relative to plots if Figure 4
%   -- major revision of code -- > lots of repetition, hard to follow
%   -- Figure 2: add * against traits used in meta-analysis

@@

\mstype{Research Article}
\runninghead{Influence of traits on growth changes with size}
\keywords{maximum height, photosynthetic rate, ontogenetic stage, functional trait, growth strategy, seed mass, specific leaf area, wood density}

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary 268/350}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate is seen as a central element of its ecological strategy, and as determined by its traits. Yet the literature is inconsistent about the empirical correlation between functional traits and growth, casting doubt on the capacity of some prominent traits to influence growth rate.

\item We propose that traits should influence growth in a way that depends on the size of individual plants. We outline mechanisms and hypotheses based on new theory, and test these predictions in tree species using a meta-analysis of 103 studies ($>$ 500 correlations) for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate). To capture the effects of size, we tested for a shift in the direction of correlation between growth rates and traits across three plant stages: seedling, sapling and adult.

\item Results were consistent with predictions. Specific leaf area was correlated with relative growth rate (RGR) in seedlings but not in adult plants. Correlations of RGR with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with RGR only in one plant stage category: seedlings and adults, respectively. Yet, despite a sizeable literature on traits and growth, coverage was still limited for wood density, seed mass, maximum height and assimilation rate. Additional studies are needed for some trait-growth-stage relationships.

@@

The phrase ``plant ecological strategy'' refers to how species face common challenges of acquiring sufficient water, nutrients, and light for growth; and then ultimately replacing themselves with offspring. The idea that a strategy differences among species might involve fast versus slow growth goes back at least to \citet{Grime:1975gr}, who observed a wide spread of relative growth rate (RGR) among seedlings growing under favourable conditions. Seedling potential RGR has continued to be used as a major strategy indicator within Grime's CSR (Competitor -- Stress tolerator -- Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling RGR into different components showed that specific leaf area (SLA) was typically the strongest source of variation between species \citep{Poorter:1989tx, Rees:2010gk}. From the 1990s a new style of ``trait ecology'' arose whereby measurable traits, including SLA, were used directly as strategy axes \citep{Westoby:2002ft}. An approach via measurable traits made broad comparisons possible across continents, latitudes, and thousands of species. Since then, traits have been seen as constructive approaches to understanding plant growth strategies \citep{Westoby:2002ft}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

Currently, two main spectra of trait variation are widely cited as underpinning differences between species in vegetative growth rates. The ``leaf economic spectrum'' \citep{Wright:2004jb} and the ``wood economic spectrum'' \citep{Chave:2009iy} reflect tissue construction costs (for leaf area and for wood volume respectively), trading off with rates of tissue turnover or mortality. It has been widely expected that species with low tissue construction cost will typically have fast growth rates, at least in favourable physical environments \citep[e.g.][]{MullerLandau:2004dc,Poorter:2008iu,Chave:2009iy,Larjavaara:2010bn,Iida:2012jb,Paine:2015df}. Yet, recently a number of empirical studies indicate the correlation between traits and growth is not as consistent as that. Consider SLA as an illustration. By definition, species with high SLA can deploy more leaf area for a given dry mass investment in leaves \citep{Poorter:1999wd, Reich:1992wm}. In addition, SLA is correlated with specific photosynthetic capacity per unit of leaf mass and high nitrogen or phosphorus concentrations in the leaf \citep{Wright:2004jb}. For seedling RGR under favourable conditions, SLA has repeatedly been found to be a strong predictor regardless of vegetation type or growth form \citep{Lambers:1992bj,Reich:1992wm,Grime:1997wm,Poorter:1999wd,Wright:1999ds}. Yet for adult plants, the correlation has mostly been much weaker or absent \citep{coomes_comparison_1998,Poorter:2008iu,Aiba:2009ft,Easdale:2009gv,Wright:2010tp}. The discrepancy between theoretical expectation and empirical results has led some to question whether traits are even useful for understanding growth \citep{Wright:2010tp, Paine:2015df}.

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Size has long been known to play a central role in organismal design and function \citep{Vogel:1988ux, Vogel:2003wb}, including that of trees \citep{Farnsworth:1995im, Givnish:1995ta}. Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may in fact be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii}. As plants grow larger, a greater amount of the energy is devoted towards building and maintaining non-photosynthetic tissues \citep{Givnish:1995ta}. A model by \citet{Falster:2011ii} suggests this change shifts the benefits and costs of some trait-based trade-offs, such that the net effect on the trait on growth changes with size. While recent empirical studies support the idea that the relationship between traits and growth rates may change with size (\citealt{Iida:2014ep, Iida:2014hq}), the trend and mechanisms behind this variation remain unclear.

In this paper, we investigate for five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass):
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from diverse studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with effects of plant size, they have generally reported trait-growth correlations for identifiable stage classes (i.e. seedling, sapling, and adult). We therefore used plant stage as a surrogate for the effects of plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms. We chose to assess hypotheses via meta-analysis because there already exist a considerable number of studies with salient information, allowing us to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experiential study. While, we anticipated a reasonably-sizeable literature on trait and growth to be available, we were uncertain about the representation of all trait-stages combinations. An additional goal of our study was therefore to identify any gaps in knowledge needed to improve our understanding of the link between trait and growth.

@@

Recent theoretical work \citep{Falster:2011ii} allows us to formulate hypotheses about relationships between traits and growth, based on a mathematical decomposition of plant growth rate. Specific predictions arise from considering the influence of traits on elements of these equations and how these effects may vary with size (see Appendix \ref{sec:growth} for mathematical explanations). This model extends earlier efforts to model growth as a function of traits \citep{Lambers:1992bj,Cornelissen:1996hf,Wright:2000kw,Enquist:2007ek}. However, whereas earlier models focused solely on mass-based RGR and did not predict a change in effect of traits with size, the revised equations do allow for these effects, while also making predictions for both AGR and RGR in any of height, diameter or mass.

We adapted growth equations from \cite{Falster:2011ii} to generate hypotheses on both absolute and relative growth rate in mass (eqn \ref{eq:eq1}), height (eqn \ref{eq:eq2}) and diameter (eqn \ref{eq:eq4}).

\begin{equation}\label{eq:eq0}
\underbrace{\frac{db}{dt}}_{\text{net live biomass production}} = \underbrace{y}_{\text{yield}}  \big(\underbrace{a_{l} A_{\textrm{area}}}_{\text{photosynthesis}} - \underbrace{\sum_{i = l,s,r,f} m_{i} r_{i}}_{\text{respiration}}\big) - \underbrace{ \sum_{i = l,s,r,f} m_{i} k_{i}}_{\text{turnover}},
\end{equation}
where $y$ is the ratio of carbon fixed in mass per carbon assimilated, and $A_{\textrm{area}}$ is the assimilation rate of CO$_{2}$ per leaf area.

\begin{equation}\label{eq:eq1}
\underbrace{{\frac{dm_t}{dt}}}_{\text{growth rate in mass}} = \underbrace{\frac{dm_t}{db}}_{\text{allocation to vegetative growth}}  \times \underbrace{\frac{db}{dt}}_{\text{net live biomass production}},
\end{equation}
where allocation to growth is the fraction of net live biomass increase that goes to vegetative growth rather than to reproductive tissue production.

The \textbf{growth rate in height} can then be expressed as:
\begin{equation}\label{eq:eq2}
\underbrace{\frac{dH}{dt}}_{\text{growth rate in height}} = \underbrace{\frac{dH}{da_{l}}}_{\text{architecture layout}} \times \underbrace{{\frac{da_{l}}{dm_{t}}}}_{\text{leaf deployment per mass}} \times {\frac{dm_t}{db}} \times {\frac{db}{dt}}.
\end{equation}
Similarly \textbf{growth rate in basal area} can be expressed as:
\begin{equation}\label{eq:eq3}
\underbrace{{\frac{da_{st}}{dt}}}_{\text{basal area growth rate}} = \underbrace{\frac{da_{ss}}{da_{l}}}_{\text{sapwood area per leaf area}} \times {\frac{da_{l}}{dm_{t}}} \times  {\frac{dm_t}{db}} \times {\frac{db}{dt}} +  \underbrace{ \frac{da_{st}}{dt}}_{\text{sapwood turnover}}.
\end{equation}

For the case of seedlings, eqs. \ref{eq:eq0}-\ref{eq:eq1} can be simplified considerably. Turnover of all tissues and allocation to reproduction have not yet begun. It is also common to assume all respiration rates are proportional to leaf area. Biomass growth rate in seedling then becomes a linear function of leaf area:
\begin{equation}\label{eq:eq1_seedlings}
{\frac{dm_t}{dt}}  \approx  a_{l} \, y \, (A_{\textrm{area}} - r_{l}).
\end{equation}

@@

Plants with high SLA can deploy more leaf area per unit of mass invested. However, high-SLA species also have shorter leaf lifespan, and must use more energy replacing leaves lost via turnover \citep{Wright:2004jb}. Whether plant growth rate increases with SLA depends on the how the relative magnitude of the benefit (cheap leaf construction) and cost (higher leaf turnover) of increased SLA also change with size (Table \ref{tab:trade-offs}). During early life, when leaf comprises a large part of the plant, decreasing SLA has an overwhelmingly positive effect on growth rate. As plants increase in size and leaf mass fraction decreases, the benefit of cheaper leaf construction diminishes; thus the cost of replacing leaf turnover becomes more important.

For seedlings (eq. \ref{eq:eq1_seedlings_RGR}), cheaper leaf construction (high SLA) should translate directly into faster relative growth rate in mass. For larger plants (eq. \ref{eq:eq1}-\ref{eq:eq4}), SLA is expected to affect growth rate positively through more cost-efficient deployment of leaf area (via $da_l/dm_t$; the same effect as in seedlings), but also negatively via increased leaf turnover (eqn 1; Table \ref{tab:trade-offs}). The effect of SLA on $da_l/dm_t$ decreases as plant size increases and increasing amounts of the plant are wood. Consequently the effect of turnover costs in slowing growth comes to dominate at larger sizes.

@@

High WD increases the cost of building stem tissue and thus the cost of deploying additional leaf area (Table \ref{tab:trade-offs}), which should slow growth. Although expensive wood decreases the risk of mortality, this benefit is not reflected in growth outcomes. Thus increasing WD is predicted to decrease growth across the entire size range.

Cheaper stem construction (low WD or stem density) increases the marginal rate of leaf deployment per total mass increment ($da_l/dm_t$), and consequently increases growth rate in leaf area, height, and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}; Table \ref{tab:trade-offs}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations. Under this scenario, wood density is negatively correlated to growth rate across all plant stages. Because the relative expenditure on stem compared to leaf is increases as plants grow larger, the relationship between WD and growth may also strengthen in larger plants.

\textbf{H3: The relationship between Hmax and growth rate is expected to be absent in seedlings but to strengthen with increasing plant size}

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do,Thomas-2011, Wenk:2015jz}). For seedlings, where all individuals are investing solely in growth, we expect no influence of Hmax on growth. As plants become larger, species with smaller Hmax are expected to be allocating more to reproduction and less to growth, because at any given size they are closer to their Hmax (Table \ref{tab:trade-offs}).

Hmax affects growth via the fraction of energy allocated to vegetative growth versus to seed production ($dm_t/db$ in eq. \ref{eq:eq1}; Table \ref{tab:trade-offs}). For seedlings and seedlings this fraction is 1 for all species and so no effect is expected. But once some species begin to produce seed, then species closer to their Hmax value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller Hmax are expected to show slower vegetative growth rates.

\textbf{H4: The relationship between seed mass and RGR is negative in seedlings, but weakens with increasing size}

The main effect of increased seed size is to produce a larger overall seedling. Under similar light conditions, RGR decreases with size because the ratio of leaf area to total biomass decreases with increasing plant size. Thus, larger seedlings are predicted to have a lower RGR during the earliest phases of growth \citep{Turnbull:2012ew}. However, among larger plants we expect the effect of this initial size to be minimal, and thus the correlation between seed size and growth rate should disappear.

The $\textrm{RGR}_{\textrm{mass}}$ slows as plants grow larger, and large-seeded species begin larger as seedlings. For larger plants, seed mass is not expected to have any distinct influence on growth rates  (Table \ref{tab:trade-offs}).
\textbf{H5: The relationship between Aarea and growth rate is expected to be positive across all plant sizes, at least under high-light conditions}

@@

The effects of benefits (higher assimilation) and costs (higher respiration) of increased Area occur within the same element of eq. \ref{eq:eq0}, and assimilation must outweigh respiration for there to be growth, so the net effect will be positive (Table \ref{tab:trade-offs}).

Note that the predicted relationship between growth and either SLA or seed mass in seedlings differs if AGR in mass is used instead of RGR$_\textrm{mass}$ (i.e. becomes negative for SLA, positive for seed mass). In our study, all data recorded for these trait-stage combinations have been measured on a RGR base for mass (see Fig S9).

\textbf{Effect on absolute and relative growth rates}
Tree growth in diameter, height or mass can be expressed as either absolute growth rate (AGR) or relative growth rate (RGR). The choice of the growth measurement is generally confounded with the size of the plant; RGR in mass is mainly used for seedling, and AGR in diameter or height are mainly used for adult. According to the growth model, the predicted relationship between growth and seed mass in seedlings reverses if AGR$_\textrm{mass}$ is used instead of RGR$_\textrm{mass}$ (see Appendix \ref{sec:growth}). The correlation with RGR$_\textrm{mass}$ is negative but becomes positive with AGR$_\textrm{mass}$. For SLA, WD, Hmax, and Aarea, the direction of the predicted relationship is not affected by the growth measurement used. However RGR and AGR are fundamentally different for a lot of ecologists. So we adopted a two steps approach, first we were conservative by focusing exclusively on RGR, then we compared results with those of a more inclusive dataset with both RGR and AGR.

@@

We aimed to identify studies reporting the correlations, measured across species, between functional traits and one of several measures of growth  rate (Table \ref{tab:trade-offs}) . We supplemented records from our own bibliographic libraries with results from searching online databases and suggestions from other researchers. Systematic searches were conducted across the two most comprehensive bibliographic platforms: Google scholar and Web of Science (Thomson Reuters). While Web of Science guarantees a relatively stable and reproducible search environment, with clearly defined lists of indexed journals, its searches are limited to meta-data (i.e. title, keywords, and abstracts, \citealt{Beckmann:2012hn}). This limitation does not apply to Google Scholar, where all text that is electronically available on the internet is indexed.

The biggest challenge when searching literature databases is to identify suitable studies from the vast corpus available. Initially we were interested to know whether data existed for a variety of traits; however the number of potential search terms made systemic searching for all these terms in Google Scholar impractical. A preliminary investigation of search results for a variety of traits on the Web of Science platform suggested that searching only on SLA and WD -- two of the most widely measured traits -- would capture  $\sim$80\% of search records that would be returned by searching for a wider range of traits (see Appendix \ref{app:literature-snapshot}). We therefore narrowed our searching to only cover these two main traits.

Both Web of Science and Google Scholar databases were searched for records linking SLA and WD to growth rate. On Web of Science, we searched using  search terms: (``trait'' AND (``growth rate'' OR ``relative growth rate'')) AND (``tree*'' OR ``woody*''), where ``trait'' was either (``leaf mass per area'' OR ``specific leaf area'' OR ``leaf construction cost'') or (``wood density'' OR``WD'' OR ``wood specific gravity'' OR ``stem density'' OR ``stem specific density'' OR ``SSD'' ).  On Google scholar we searched using terms: ((``SLA'' OR ``LMA'' OR ``specific leaf area'' OR ``leaf mass area'' OR ``wood density'' OR ``WD'' OR ``wood specific gravity'' OR ``stem density'' OR ``stem specific density'') AND (``growth rate'' OR ``relative growth rate'')) AND (``tree*'' OR ``woody*''). This keyword search in Google Scholar returned approximately 18900 papers. We examined all records returned from the Web of Science search (XXX) and the first 750 items from Google Scholar when ranked by ``relevance''. All searches were conducted between March and December 2014.

In total we inspected the full text of 583 articles (see Appendix XXXX in Supplementary material for list). Individual studies had to report the following information to be included in our dataset: i) the relationship between growth and trait as a correlation coefficient, ii) the sample size and the number of species used to perform the correlation, and iii) the size or stage of plants. Despite searching only on SLA and WD, we recorded coefficient correlations for a list of 19 traits in total: the five for which we had prior expectations, plus any of 14 others commonly used in the trait literature. Where available, we also recorded the geographic coordinates, number of sites, experiment type (greenhouse, field experiment, forest, shade-house etc.), growth measure (relative or AGR measured using plant diameter, height or mass), vegetation type, life form (tree, woody or mixed with herbaceous species etc.), biological scale (interspecific, intraspecific or intragenus), and experimental treatments (e.g. fertilization, shade, salt stress). In some cases, we asked the authors for additional information about the specific correlation coefficient value when this was not already provided.

We recorded coefficient correlations for a list of 19 traits (Fig. \ref{fig:fig1}): the five for which we had prior expectations, plus 14 others commonly used in the trait literature. SLA, leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and Aarea are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. Seed mass and Hmax capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. WD and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here WD and SA/LA are also considered as indicators of stem construction cost \citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource distribution to leaf \citep{Lambers:1992bj}.
